### 2018-01-11 manuscript.rmd
expanded broader perspectives; conclusions

New data were analysed in conjunction with 13 publicly available datasets, with a total number of 43 species. Sequence assembly, annotation, homology evaluation, gene tree construction, parsing of genes trees to isolate orthologous sequences, and supermatrix construction were conducted with Agalma v. 1.00 [@Dunn:2013kw; @Guang202416].  This workflow integrates a variety of existing tools [@grabherr2011full, @altschul1990basic, @doi:10.1093/nar/30.7.1575, @stamatakis2006raxml, @sukumaran2010dendropy, @langmead2012fast, @li2011rsem, @talavera2007improvement, @li2009sequence, @katoh2013mafft] and new methods. Maximum likelihood analyses of the supermatrix were conducted with [XX add raxml version and citation]. Bayesian Inference (BI) analyses of the supermatrix were conducted using Phylobayes v. 1.7a-mpi [@lartillot2009phylobayes]. Sequence alignments, sampled and consensus trees, and voucher information are available in the git repository. Tree figures were rendered with ggtree [@Yu:2016fo].

@@

Most relationships received strong support across analysis methods (Fig. 3A), with a couple localized exceptions (Fig. 3B). The phylogenetic relationships recovered in this study are largely consistent with those found in a previous study based on two genes (16S and 18S ribosomal RNA) [@Dunn:2005dy]. Relationships that receive strong support in both studies include the placement of Cystonectae as sister to Codonophora (the clade that includes all other siphonophores), the placement of Apolemiidae as sister to all other codonophorans, and the placement of Calycophorae within the paraphyletic "Physonectae". Multiple nodes that were not resolved in the previous two-gene analysis receive strong support in the present 1,071-gene transcriptome analyses. There is strong support for the Pyrostephidae as sister to all other non-apolemiid codonophorans. Within the clade that is sister to the Pyrostephidae, we find two main clades, the Calycophorae and a clade we here name Euphysonectae (Fig. 3A). It includes the remaining non-apolemiid, non-pyrostephid "Physonectae". We define the Euphysonectae as the clade consisting of *Agalma elegans* and all taxa that are more closely related to it than to *Diphyes dispar*.

@@

Within Clade A, *Physophora gilmeri* along with *Lychnagalma utricularia* (both not included in the previous phylogeny) are sister to the Agalmatidae, a clade restricted to *Agalma*, *Athorybia*,*Melophysa*, *Halistemma* and *Nanomia* [@Dunn:2005dy; @pugh2006taxonomic]. In the rDNA study, *Physophora hydrostatica* (the presumed sister species to *P. gilmeri*) was sister to the Forskaliidae with low support. The position of *Cordagalma cordiforme* (= *Cordagalma ordinatum*) [@pugh2016synopsis] was previously unresolved, while in this analysis *Cordagalma* sp. is in a clade with *Forskalia asymmetrica*, falling outside of the Agalmatidae. Placement of *Cordagalma* outside Agalmatidae is consistent with previous analyses of morphological and molecular data [@pugh2006taxonomic; @Dunn:2005dy].

@@

@@

Ancestral character state reconstructions favor the hypothesis that monoecy arose twice (Fig. 5A), once in Calycophorae and once in Clade A. This is consistent with differences in the arrangements of male and female gonophores in the two clades. In Clade A, male and female zooids are found within the same cormidium. In these species, the male and female zooids are placed at very different but well defined locations within the cormidium. Meanwhile in calycophorans, each cormidium (a single reiterated sequence of zooids along the stem, see Fig. 2) bears either male or female zooids. In this form of monoecy, the male and female cormidia can either be in an alternating pattern, or there can be several male or female cormidia in a row. In either case, male and female zooids are found at the same corresponding locations within the cormidia. However, there are exceptions -- in *Sulculeolaria*, colonies appear to consist of only one sex, however they are monoecious and protoandrous, with female gonophores developing after the release of male gonophores [@carre1979].

@@

Bracts are highly reduced zooids unique to siphonophores, where they are only present in Codonophora (Fig. 4). Bracts are functional for protection of the delicate zooids and to help maintain neutral buoyancy [@jacobs1937beobachtungen]. Some calycophorans are able to actively exclude sulphate ions in their bracts to adjust their buoyancy along the colony [@bidigare1980role]. The common ancestor of Codonophora had only one bract subtype [XX provide citation, indicate which analyses support this, or remove]. Bracts were lost in Hippopodidae, some clausophyids, *Physophora hydrostatica*, and in *Gymnopraia lapislazula*. These patterns of loss are not captured by the present sampling, as most of these species are not included in the present phylogeny.

@@

Siphonophores are abundant predators in the pelagic realm, ranging from the surface (*Physalia physalis*) to bathypelagic depths (Figs. 4 and S9) [@mapstone2014global; @mackie1987siphonophore]. The depth distribution of siphonophore populations is not always static, as some species are known to be vertical migrators, although this is within a relatively narrow depth range (<100m) [@pugh1984diel]. Some species such as *Nanomia bijuga* exhibit synchronous diel migration patterns [@barham1966deep]. Using the present phylogeny, we reconstructed the median depth changes along the phylogeny under a Brownian Motion model (Fig. S9), which had the strongest AICc support (compared to non-phylogenetic distributions, and to Ohrnstein-Uhlenbeck). This model indicates a mesopelagic most recent common ancestor, with several independent transition events to epipelagic (light blue) and bathypelagic (dark blue) waters. This reconstruction only included species recorded using an ROV, thus it excludes many other independent colonizations of the epipelagic habitat. The ROV observations are reliable below 200m, and no quantative measurements were made on SCUBA dives. Species such as *Hippopodius hippopus*, *Athorybia rosacea*, *Diphyes dispar*, and *Chelophyes appendiculata* are often encountered blue water diving less than 20m from the surface (Fig. 4). We also reconstructed the median depth changes along the phylogeny using median depths of 20m for all species collected by SCUBA diving (Fig. S10), and still find support for a mesopelagic ancestor. There was only a single transition to benthic lifestyle on the branch of *Rhodaliidae*, and a single transition to a pleustonic lifestyle on the branch of *Physalia physalis*.

## Conclusions
Using phylogenomic tools we were able to resolve deep relationships within Siphonophora with strong support. A previous rRNA study was unable to resolve deep relationships within Codonophora, while our study suggests that Pyrostephidae are sister to all other non-apolemiid codonophorans. Our study supports the rejection of the monophyly of Physonectae, a traditional morphological grouping, and rejects the monophyly of monoecious siphonophores. Future studies with the addition of key taxa within Codonophora will enable resolution of internal relationships within Siphonophora, and shed light on patterns of character evolution that cannot be fully assessed here.

>>>>

New data were analysed in conjunction with 13 publicly available datasets, with a total number of 43 species. Sequence assembly, annotation, homology evaluation, gene tree construction, parsing of genes trees to isolate orthologous sequences, and supermatrix construction were conducted with Agalma v. 1.00 [@Dunn:2013kw; @Guang202416].  This workflow integrates a variety of existing tools [@grabherr2011full, @altschul1990basic, @doi:10.1093/nar/30.7.1575,, @sukumaran2010dendropy, @langmead2012fast, @li2011rsem, @talavera2007improvement, @li2009sequence, @katoh2013mafft] and new methods. Maximum likelihood analyses of the supermatrix were conducted with RAxML v 8.2.0 [@stamatakis2006raxml] and implemented via Agalma. Bayesian Inference (BI) analyses of the supermatrix were conducted using Phylobayes v. 1.7a-mpi [@lartillot2009phylobayes]. Sequence alignments, sampled and consensus trees, and voucher information are available in the git repository. Tree figures were rendered with ggtree [@Yu:2016fo].

@@

Most relationships received strong support across analysis methods (Fig. 3A), with a couple of localized exceptions (Fig. 3B). The phylogenetic relationships recovered in this study are largely consistent with those found in a previous study based on two genes (16S and 18S ribosomal RNA) [@Dunn:2005dy]. Relationships that receive strong support in both studies include the placement of Cystonectae as sister to Codonophora (the clade that includes all other siphonophores), the placement of Apolemiidae as sister to all other codonophorans, and the placement of Calycophorae within the paraphyletic "Physonectae". Multiple nodes that were not resolved in the previous two-gene analysis receive strong support in the present 1,071-gene transcriptome analyses. There is strong support for Pyrostephidae as sister to all other non-apolemiid codonophorans. Within the clade that is sister to the Pyrostephidae, we find two main clades, Calycophorae and a clade we here name Euphysonectae (Fig. 3A). It includes the remaining non-apolemiid, non-pyrostephid "Physonectae". We define Euphysonectae as the clade consisting of *Agalma elegans* and all taxa that are more closely related to it than to *Diphyes dispar*.

@@

Within Clade A, *Physophora gilmeri* along with *Lychnagalma utricularia* (both not included in the previous phylogeny) are sister to the Agalmatidae, a clade restricted to *Agalma*, *Athorybia*,*Melophysa*, *Halistemma* and *Nanomia* [@Dunn:2005dy; @pugh2006taxonomic]. In the rDNA study, *P. hydrostatica* (the presumed sister species to *P. gilmeri*) was sister to the Forskaliidae with low support. The position of *Cordagalma cordiforme* (= *Cordagalma ordinatum*) [@pugh2016synopsis] was previously unresolved, while in this analysis *Cordagalma* sp. is in a clade with *Forskalia asymmetrica*, falling outside of the Agalmatidae. Placement of *Cordagalma* outside Agalmatidae is consistent with previous analyses of morphological and molecular data [@pugh2006taxonomic; @Dunn:2005dy].

@@

The broad sampling approach of this phylogeny provides new evidence for the relationships between major siphonophore clades within Codonophora, specifically between Pyrostephidae, Calycophorae, and the newly named Euphysonectae. This opens up new questions about key relationships within both Calycophorae and Euphysonectae -- where future transcriptome sampling efforts should be focussed. Within Euphysonectae, two clades (Clade A and Clade B) are hypothesized, however there is weaker support for Clade B (Fig. 3A, 3B). Expanding sampling of species that probably fall in Clade B, including other *Erenna* species, rhodaliids, and relatives of Undescribed sp L, will greatly expand our understanding of these two groups and perhaps provide evidence of Clade B synapomorphies. Similarly, within Calycophorae, increased taxon sampling is needed. This study, and the previous study, suggest that the prayomorphs are paraphyletic, but for slightly different reasons. In @Dunn:2005dy, a clade of prayomorphs including *Praya dubia*, *Nectadamas diomedeae*, and *Nectopyramis natans* (not included in this study) were found to be sister to all other calycophorans, while in this study, the prayomorph *Lilyopsis flyoracantha* (not included in the previous study) is found in a clade including diphyomorph calycophorans that is sister to all other prayomorphs. Expanded taxon sampling, particularly *P. dubia* or a nectopyramid, but also extensive sampling across the major prayomorph and diphyomorph groups, will expand our understanding of relationships within Calycophorae. This will be especially important for understanding trait evolution within Calycophorae,  for example, the release of eudoxids (Fig. 4), or the arrangement of male and female zooids along the stem (see below).

@@

Ancestral character state reconstructions favor the hypothesis that monoecy arose twice (Fig. 5A), once in Calycophorae and once in Clade A. This is consistent with differences in the arrangements of male and female gonophores in the two clades. In Clade A, male and female zooids are found within the same cormidium. In these species, the male and female zooids are placed at very different but well defined locations within the cormidium. Meanwhile in calycophorans, each cormidium (a single reiterated sequence of zooids along the stem, see Fig. 2) bears either male or female zooids. In this form of monoecy, the male and female cormidia can either be in an alternating pattern, or there can be several male or female cormidia in a row. In either case, male and female zooids are found at the same corresponding locations within the cormidia. However, there are exceptions -- in *Sulculeolaria* (not included in this phylogeny), colonies appear to consist of only one sex, however they are monoecious and protoandrous, with female gonophores developing after the release of male gonophores [@carre1979].

@@

Bracts are highly reduced zooids unique to siphonophores, where they are only present in Codonophora (Fig. 4). Bracts are functional for protection of the delicate zooids and to help maintain neutral buoyancy [@jacobs1937beobachtungen]. Some calycophorans are able to actively exclude sulphate ions in their bracts to adjust their buoyancy along the colony [@bidigare1980role]. Bracts were lost in Hippopodidae, some clausophyids, *Physophora hydrostatica*, and in *Gymnopraia lapislazula*. These patterns of loss are not captured by the present sampling, as most of these species are not included in the present phylogeny.

@@

Siphonophores are abundant predators in the pelagic realm, ranging from the surface (*Physalia physalis*) to bathypelagic depths (Figs. 4 and S9) [@mapstone2014global; @mackie1987siphonophore]. The depth distribution of siphonophore populations is not always static, as some species are known to be vertical migrators, although this is within a relatively narrow depth range (<100m) [@pugh1984diel]. Some species such as *Nanomia bijuga* exhibit synchronous diel migration patterns [@barham1966deep]. Using the present phylogeny, we reconstructed the median depth changes along the phylogeny under a Brownian Motion model (Fig. S9), which had the strongest AICc support (compared to non-phylogenetic distributions, and to Ohrnstein-Uhlenbeck). This model indicates a mesopelagic most recent common ancestor, with several independent transition events to epipelagic (light blue) and bathypelagic (dark blue) waters. There was only a single transition to benthic lifestyle on the branch of *Rhodaliidae*, and a single transition to a pleustonic lifestyle on the branch of *Physalia physalis*. Depth appears to be phylogenetically conserved in Euphysonectae after the split between Clade A (shallow living species) and Clade B (deep dwelling species); however several shallow-living species that likely belong in Clade B were not included in this analysis, suggesting that depth may be more labile than it appears. Phylogenetic conservatism is also observed for the mesopelagic Apolemiidae and Pyrostephidae clades, but not for the Calycophorae, which show a diversified evolutionary trajectory for depth. The present sampling, however, is not sufficient to capture significant variation in depth distributions between closely related species. Previous studies have shown that many species that are collected at the same locality are found to occupy discrete, largely non-overlapping depth distributions, including between species that are closely related [@pugh1974vertical].

This reconstruction (Fig. S9) only included species recorded using an ROV, thus it excludes many other independent colonizations of the epipelagic habitat. The ROV observations are reliable below 200m, and no quantative measurements were made on SCUBA dives. Species such as *Hippopodius hippopus*, *Athorybia rosacea*, *Diphyes dispar*, and *Chelophyes appendiculata* are often encountered blue water diving less than 20m from the surface (Fig. 4). We also reconstructed the median depth changes along the phylogeny using median depths of 20m for all species collected by SCUBA diving or swimming (Fig. S10), and still find support for a mesopelagic ancestor. It is important to note, however, that *H. hippopus* and *C. appendiculata* were both collected in the bay of  Villefrance-sur-mer, France, where an upwelling is known to bring deeper species closer to the surface [@nival1976evolution].

## Conclusions
Using phylogenomic tools we were able to resolve deep relationships within Siphonophora with strong support. A previous rRNA study was unable to resolve deep relationships within Codonophora, while our study suggests that Pyrostephidae are sister to all other non-apolemiid codonophorans. We find support for two monophyletic groups within the clade that is sister to Pyrostephidae -- Calycophorae and a new clade we name Euphysonectae. Within Euphysonectae, we find support for two monophyletic groups dubbed Clade A and Clade B. Our study supports the rejection of the monophyly of Physonectae, a traditional morphological grouping, and rejects the monophyly of monoecious siphonophores. These results are consistent with a hypothesis that monoecy evolved twice, once in Clade A and once in Calycophorae. This hypothesis is consistent with differences in the arrangement of male and female zooids in the two monoecious groups. Future studies with the addition of key taxa within Euphysonectae will enable resolution of internal relationships within Siphonophora, and shed light on patterns of character evolution, including the evolution of zooid types, that cannot be fully assessed here.
